---
title: "Map Making in R: An Example"
author: "Anna Talucci"
date: "04/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

A working example of assembling a map with `ggplot()` and `cowplot()`.  The use of `ggplot()` for mapping is dependent on the use of `sf` objects

- This script is reliant on relative path names "../folder-name/file-name.ext"
- Please reference the Git Hub repository called **BC-Maps** to see the file  structure  used here

# Packages

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(sf)
library(ggmap) # devtools::install_github("dkahle/ggmap")
library(ggrepel)
library(raster)
library(rgdal)
library(rasterVis)
library(RColorBrewer)
library(cowplot)
library(ggspatial)
library(maps)
```

# Read in shapefile boundaries

We are using relative file path names here. Note the file name is preceded by the sub-folder shapefiles in the data folder.

```{r}
fire1 <- st_read(
  "../shapefiles/ch_R10070_2014.shp")
```

```{r}
fire2 <- st_read(
  "../shapefiles/en_R10171_2012.shp")
```

```{r}
fire3 <- st_read(
  "../shapefiles/tw_R10252_2013.shp")
```

```{r}
all_fires= st_read("../shapefiles/perimeter_merge.shp")
```

```{r}
bc = st_read("../shapefiles/province_bc_alber.shp")
```

```{r}
lakes = st_read("../shapefiles/Lakes_select_fire.shp")
```

```{r}
parks = st_read("../shapefiles/tweed-entiako_singlepoly.shp")
```

```{r}
city = st_read("../shapefiles/BC_Vancouver_PrinceGeorge.shp")
```

# Create a bounding box 

Use metadata from all_fires shapefile to identify `xmin`, `xmax`, `ymin`, and `ymax` to define the bounding box

```{r}
(insetrect <- data.frame(xmin = 985362, xmax = 1069051, ymin = 894097.7, ymax = 939814.2))
```

# Map Making

Code annotations are preceded by `#` and precede the line of code they describe.

## Insert city location with name label

Use `str()` on the city shapfile to look at the data frame for the point file to identify data frame components to use in `ggplot()` e.g. lat, lon for coordinates  and NAME for city name label.

```{r}
ggplot() + 
  #add BC Boundary
  geom_sf(data = bc, fill = "gray75", color = "black") + 
  #add city point data
  geom_point(data = city, aes(x = lon, y= lat), shape = 24, size = 2, color = "red", fill = "red") + 
  #add Label city name
  geom_label_repel(data = city, aes(x = lon, y= lat, label = c("Prince George", "Vancouver")), vjust = "top", hjust = "left", color = "black") + 
  xlab("")+ylab("") +
  coord_sf()
```

## British Columbia with study area bounding box

This map will serve as an inset in the final map.

```{r}
map_bc2 = ggplot() + 
  geom_sf(data = bc, fill = "#E8E8E8", color = "black") +
  #Use bounding box from above
  geom_rect(data = insetrect, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), alpha = 0, colour = "#000080", size = 1.5, linetype = 1) +
  xlab("")+ylab("")+
  coord_sf() +
   theme_void() +
  #add a bounding box so that will border the inset
  theme(panel.background = element_rect(colour = "black", fill = "white", size = 0.5))

map_bc2
```

## Map with legend bottom

Here e create a map with a legend box that is placed outside of the map along the bottom.

```{r}
map_legend2 = ggplot() +
  #add shapefile
  geom_sf(data = bc, aes(fill = "BC", color = "BC")) + 
  #add shapefile
  geom_sf(data = parks, aes(fill = "Parks", color = "Parks"), show.legend = TRUE) + 
  #add shapefile
  geom_sf(data = lakes, aes(fill = "Lakes", color = "Lakes"), show.legend = TRUE, key_glyph = draw_key_rect) + 
  #add shapefile
  geom_sf(data = fire1, fill = NA, aes(color = "Fire1"), size = 1, show.legend = TRUE, key_glyph = draw_key_rect) + 
  #add shapefile
  geom_sf(data = fire2, fill = NA, aes(color = "Fire2"), size = 1, show.legend = TRUE)  +
  #add shapefile
  geom_sf(data = fire3, fill = NA, aes(color = "Fire3"), size = 1, show.legend = TRUE) + 
  #manually define colors with hexcodes 
  scale_color_manual(values = c("BC" = "#E8E8E8", "Parks" = "#C3ECB2", "Lakes" = "#AADAFF", "Fire1" = "#b82b2b", "Fire2" = "#b06f0c", "Fire3" =  "#dd7116"), 
                       labels = c("British Columbia", "Chelaslie fire", "Entiako fire", "Tweedsmuir fire", "Lakes", "Provincial parks"),
                       name = "") + 
  #manual define fill; for no fill NA
  scale_fill_manual(values = c("BC" = "#E8E8E8", "Parks" = "#C3ECB2", "Lakes" = "#AADAFF", "Fire1" = NA, "Fire2" = NA, "Fire3" =  NA), guide=FALSE) + 
  #define coordinates based on shapefile metadata
  coord_sf(crs = "+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs", xlim = c(983000, 1071000), ylim = c(892000, 941000), expand = FALSE)  + 
  #add a scale bar
  annotation_scale(location = "br", width_hint = 0.25, text_size = 12, text_face = NULL, text_family = "serif", text_col = "black") + 
  #add North Arrow
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.4, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_nautical(line_width = 1, line_col = "black", fill = c("black", "white"), text_size = 10, text_face = NULL, text_family = "serif", text_col = "black", text_angle = 0)) + 
  #define theme
  theme(axis.text.x = element_text(face = "plain", color = "black", size = 10),
        axis.text.y = element_text(face = "plain", color = "black", size = 10),
        legend.position= "bottom",
        legend.key = element_rect(fill = "white")) 

map_legend2
```

## Alternative legend location

Here we create a map with a legend box that is placed in the low right corner of the map.

```{r}
map_legend = ggplot() +
  geom_sf(data = bc, aes(fill = "BC", color = "BC")) +
    geom_sf(data = parks, aes(fill = "Parks", color = "Parks"), show.legend = TRUE) +
    geom_sf(data = lakes, aes(fill = "Lakes", color = "Lakes"), show.legend = TRUE, key_glyph = draw_key_rect) +
  geom_sf(data = fire1, fill = NA, aes(color = "Fire1"), size = 1, show.legend = TRUE, key_glyph = draw_key_rect) +
  geom_sf(data = fire2, fill = NA, aes(color = "Fire2"), size = 1, show.legend = TRUE) +
  geom_sf(data = fire3, fill = NA, aes(color = "Fire3"), size = 1, show.legend = TRUE) +
  scale_color_manual(values = c("BC" = "#E8E8E8", "Parks" = "#C3ECB2", "Lakes" = "#AADAFF", "Fire1" = "#b82b2b", "Fire2" = "#b06f0c", "Fire3" =  "#dd7116"), 
                       labels = c("British Columbia", "Chelaslie fire", "Entiako fire", "Tweedsmuir fire", "Lakes", "Provincial parks"),
                       name = "Legend") +
  scale_fill_manual(values = c("BC" = "#E8E8E8", "Parks" = "#C3ECB2", "Lakes" = "#AADAFF", "Fire1" = NA, "Fire2" = NA, "Fire3" =  NA),
                    guide=FALSE) +
  coord_sf(crs = "+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs", xlim = c(983000, 1071000), ylim = c(892000, 941000), expand = FALSE)  + 
      annotation_scale(location = "br", width_hint = 0.25, text_size = 12, text_face = NULL, text_family = "serif", text_col = "black") +
    annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in"),
        style = north_arrow_nautical(line_width = 1, line_col = "black", fill = c("black", "white"), text_size = 10, text_face = NULL, text_family = "serif", text_col = "black", text_angle = 0)) +
  theme(axis.text.x = element_text(face = "plain", color = "black", size = 12),
        axis.text.y = element_text(face = "plain", color = "black", size = 12),
        legend.position= c(0.88, 0.35),
        legend.key = element_rect(fill = "white")) 

map_legend
```


# Combining Maps

We combine our study area map with the inset map for a final map figure


```{r fig.height=4, fig.width=6}
fig_manuscript2 = ggdraw() +
  #add base map
  draw_plot(map_legend2) +
  #add inset and define size and location 
  draw_plot(map_bc2, x = 0.12, y = 0.32, width = .25, height = .25) 
 
fig_manuscript2
```

# Save map as file

Use `ggsave()` to export the map to a file. Note the relative path name. Multiple extensions are available. Width and height are defined for this figure, but should be adjusted for future figures.

```{r}
ggsave("../figures/bc_map_manuscript.png", plot = fig_manuscript2, width = 6, height = 4, dpi = 600)
```


